### 2021-03-08 text/draft.tex
machine indexing

In this section I will\ldots{}

@@

\hypertarget{outline}{%
\subsection{Outline}\label{outline}}

@@

\hypertarget{tools}{%
\subsection{Tools}\label{tools}}

For data refinement I use OpenRefine version 3.4.1
(https://openrefine.org/). Data manipulation in OpenRefine is tracked:
the manipulation history of some data can be exported as JSON file and
then reproduced (on the same data on a different machine or on different
data) by loading said file. I will supply a corresponding manipulation
history whenever appropriate. Note that when using OpenRefine with
larger files (and there are many such files in this project), the memory
allocation to OpenRefine must be manually increased (see
https://docs.openrefine.org/manual/installing/\#increasing-memory-allocation
for details). Also note that OpenRefine always counts the number of
records by the first column (a fact which caused me many headaches).

@@

Even though Edoc is a public server, its database does not have a
web-ready API. In addition, since Edoc is a production server, its
underlying database cannot be used directly on pain of disturbing the
provided services. The data hence needs to be extracted from Edoc in
order to work with it. This could for example be achieved by cloning the
database of the server and running it on a local machine. However, this
route was not available due to the limited resources of the responsible
co-workers. I thus employed a workaround: Edoc has a built-in advanced
search tool where the results can be exported. Even though it is not
possible to extract all database items by default, only one of the many
available search fields has to be filled in. The complete database can
hence be extracted by only filling in the \texttt{date} field and using
a sufficiently permissive time period such as 1940 to 2020 since the
oldest record in the database was published in 1942. On January 21 2021,
this query yielded 68'345 results. These results were then exported as a
326 MB JSON file called \texttt{1942-2020.json}. \texttt{1942-2020.json}
has two drawbacks. First, the maximum file size on GitHub is 100 MB. And
second, \texttt{1942-2020.json} is too big to be handled by standard
editors requiring special editors such as Oxygen. For these reasons
\texttt{1942-2020.json} was split into smaller files that are easier
handled and can be uploaded to GitHub. For this
\texttt{Utility.split\_json} was used yielding 14 files of size 20 MB or
less containing 5000 or fewer entries each. These files are called
\texttt{raw\_master\_\{year\}.json} and saved under \texttt{/files/raw}.

@@

Taking into account the third constraint simply means that we do not
employ any further restrictions. We can hence construct the sample data
set by choosing exactly those items from \texttt{/files/raw} which have
non-empty data fields \texttt{title}, \texttt{abstract},
\texttt{id\_number}, and \texttt{keywords}. We do so by calling
\texttt{Data.select\_from\_file} iteratively for all files in
\texttt{/files/raw}. The resulting file is saved in
\texttt{/files/selected} as \texttt{selected\_master.json}.

@@

\begin{figure}
\centering
\includegraphics{images/chi_square_selection_fields.pdf}
\caption{The sample data set (\(n\) = 4'111) is not representative of
the raw Edoc data per department. Data field \texttt{abstract}:
\(\chi^2 (df=9) =\) 3'160.556, \(p < 0.001\); data field
\texttt{id\_number}: \(\chi^2 (df=9) =\) 4'209.0285, \(p < 0.001\); data
field \texttt{keywords}: \(\chi^2 (df=9) =\) 2'314.533, \(p < 0.001\).
The data foundation is available at
\texttt{/files/analysis/chi\_square\_\{data\ field\}} and the
\(\chi^2\)-statistic can be calculated by calling
\texttt{Analysis.print\_chi\_square\_fit} on the data foundation files.}
\end{figure}

Of the 68'345 items in \texttt{/files/raw}, all have a title (non-empty
\texttt{title} data field), a little more than half of the items have an
abstract (37'381 items with non-empty \texttt{abstract} data field),
roughly half of the items have an ID (35'355 items with non-empty
\texttt{id\_number} data field),\footnote{Note that when using the facet
  by blank on \texttt{id\_number} in OpenRefine, there are 57'153
  matches; this is due to the fact that many items with an ID such as
  DOI have secondary or tertiary IDs such as ISI or PMID.} and less than
10\% of the items have keywords (6'660 items with non-empty
\texttt{keywords} data field). The sample data set as requires all the
above data fields to be non-empty;
\texttt{/files/sample/sample\_master.json} has 4'111 items and hence
constitutes 6\% of the raw data.

@@

\begin{figure}
\centering
\includegraphics{images/raw_sample_analysis.pdf}
\caption{The sample data set (\(n\) = 4'111) is significantly skewed
towards English (data field \texttt{language}: \(\chi^2 (df=3) =\)
1'441.414, \(p < 0.001\)) journal publications (data field
\texttt{type}: \(\chi^2 (df=10) =\) 52'519.743, \(p < 0.001\)) in the
sciences, medicine and economics (data field \texttt{department}:
\(\chi^2 (df=9) =\) 14'447.276, \(p < 0.001\)) from the 21st century
(data field \texttt{date}: \(\chi^2 (df=7) =\) 35'878.493,
\(p < 0.001\)) as compared to the raw Edoc data. The data foundation is
available at \texttt{/files/analysis/chi\_square\_\{data\ field\}} and
the \(\chi^2\)-statistic can be calculated by calling
\texttt{Analysis.print\_chi\_square\_fit} on the data foundation files.}
\end{figure}

@@

In order to generate the ouput, we call
\texttt{Data.super\_enrich\_with\_annif}, thereby iteratively calling
\texttt{Data.enrich\_with\_annif} for all items in
\texttt{/files/sample/sample\_master.json} for all Annif projects. The
output is saved in \texttt{/indexed/} as \texttt{indexed\_master.json}.

@@

As explained in \protect\hyperlink{machine-indexing}{section ``Machine
indexing''}, Annif assigns index terms from a controlled vocabulary. If
we want to assess the quality of the indexing via a gold standard, we
must therefore ensure that the gold standard makes use of the vocabulary
used by Annif. The relevant (English) vocabularies are Wikidata and YSO.
The next step in constructing the derivative gold standard is hence to
match the extracted and cleaned keywords with keywords from Wikidata and
YSO. This process is called ``reconciliation'' (see
https://docs.openrefine.org/manual/reconciling) and the tool of choice
for this task is OpenRefine (see
\protect\hyperlink{introduction}{section ``Introduction''}).

@@

In this section I will discuss the general results of assessing the
performance of the 100 Annif configurations versus the derivative gold
standard with respect to the Edoc sample data set. The data foundation
is available at \texttt{/analysis/metrics.json}.

@@

A surprising result is that there was no performance gain achieved by
increasing the text basis from titles to titles and abstracts (see
\protect\hyperlink{creating-the-data-foundation}{section ``Creating the
data foundation''}: if anything, the performance for the extended text
basis was slightly worse (see Figure 6, right). There is again no easy
explanation for this find since there are multiple plausible causes
which cannot be distangled here. For example, the abstract of an item in
the Edoc sample data set might be less representative of the item's
content than its title. In that case, the observed worse perfomance of
the Annif configurations based on title and abstracts would even be
expected. Alternatively, it might be the case that the problem lies with
Annif: a larger text basis introduces more confounding factors and might
make it harder to extract the subject terms required. An experimentum
crucis to decide between the two discussed explanations would be to
further extend the text basis to include full texts in addition to
titles and abstracts.

@@

In \protect\hyperlink{general-results}{section ``General results''} I
have identified and discussed the overall best Annif configerations.
However, in \protect\hyperlink{analysis}{section ``Analysis''}) I had
also noted the caveat that the performance of Annif might vary according
to department due to systematic biases in constructing the Edoc sample
data set. In this section I will therefore assess the performance of the
various Annif configurations per department.

@@

\hypertarget{outlook-and-conclusion}{%
\section{Outlook and conclusion}\label{outlook-and-conclusion}}

\hypertarget{refinement}{%
\subsection{Refinement}\label{refinement}}

@@

\hypertarget{implementation-strategy}{%
\subsection{Implementation strategy}\label{implementation-strategy}}

@@

Codebase available at: \url{https://github.com/MHindermann/mas}

\hypertarget{class-files.utility}{%
\subsection{class files.Utility()}\label{class-files.utility}}

A collection of utility functions.

\hypertarget{classmethod-load_jsonfile_path}{%
\subsubsection{classmethod
load\_json(file\_path)}\label{classmethod-load_jsonfile_path}}

Load a JSON object from file.

\begin{itemize}
\item
  \textbf{Parameters}

  \textbf{file\_path} (\texttt{str}) -- complete path to file including
  filename and extension
\item
  \textbf{Return type}

  \texttt{list}
\end{itemize}

\hypertarget{classmethod-save_jsondata-file_path}{%
\subsubsection{classmethod save\_json(data,
file\_path)}\label{classmethod-save_jsondata-file_path}}

Save data as JSON file.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{data} (\texttt{Union}{[}\texttt{List}, \texttt{Dict}{]}) --
    the data to be saved
  \item
    \textbf{file\_path} (\texttt{str}) -- complete path to file
    including filename and extension
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

\hypertarget{classmethod-split_jsonfile_path-save_path}{%
\subsubsection{classmethod split\_json(file\_path,
save\_path)}\label{classmethod-split_jsonfile_path-save_path}}

Split a JSON file into files not larger than 100MB.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{file\_path} (\texttt{str}) -- complete path to file
    including filename and extension
  \item
    \textbf{save\_path} (\texttt{str}) -- complete path to save folder
    including filename without extension
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

\hypertarget{class-files.data}{%
\subsection{class files.Data()}\label{class-files.data}}

A collection of Edoc data functions.

\hypertarget{classmethod-enrich_author_keywordsfile_path-save_path}{%
\subsubsection{classmethod enrich\_author\_keywords(file\_path,
save\_path)}\label{classmethod-enrich_author_keywordsfile_path-save_path}}

Enrich author keywords.

For each Edoc item: the string of author keywords is cut into single
keywords and each keyword is cleaned. Each keyword is then enriched with
Qid, MeSH and YSO ID if available.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{file\_path} (\texttt{str}) -- complete path to file
    including filename and extension
  \item
    \textbf{save\_path} (\texttt{str}) -- complete path to save folder
    including filename without extension
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

\hypertarget{classmethod-enrich_with_anniffile_path-save_path-project_ids-abstractfalse-fulltextfalse-limitnone-thresholdnone}{%
\subsubsection{classmethod enrich\_with\_annif(file\_path, save\_path,
project\_ids, abstract=False, fulltext=False, limit=None,
threshold=None)}\label{classmethod-enrich_with_anniffile_path-save_path-project_ids-abstractfalse-fulltextfalse-limitnone-thresholdnone}}

Enrich items from file with automatic keywords using Annif-client.

Available Annif-client project IDs are yso-en, yso-maui-en,
yso-bonsai-en, yso-fasttext-en, wikidata-en.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{file\_path} (\texttt{str}) -- complete path to file
    including filename and extension
  \item
    \textbf{save\_path} (\texttt{str}) -- complete path to save folder
    including filename without extension
  \item
    \textbf{project\_ids} (\texttt{List}{[}\texttt{str}{]}) --
    Annif-client project IDs to be used
  \item
    \textbf{abstract} (\texttt{bool}) -- toggle use abstract for
    indexing, defaults to False
  \item
    \textbf{fulltext} (\texttt{bool}) -- toggle use fulltext for
    indexing, defaults to False
  \item
    \textbf{limit} (\texttt{Optional}{[}\texttt{int}{]}) -- Annif-client
    limit, defaults to None
  \item
    \textbf{threshold} (\texttt{Optional}{[}\texttt{int}{]}) --
    Annif-client threshold, defaults to None
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

\hypertarget{classmethod-enrich_with_meshfile_path-save_path}{%
\subsubsection{classmethod enrich\_with\_mesh(file\_path,
save\_path)}\label{classmethod-enrich_with_meshfile_path-save_path}}

Enrich Edoc data per item with MeSH keywords from PubMed if available.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{file\_path} (\texttt{str}) -- complete path to file
    including filename and extension
  \item
    \textbf{save\_path} (\texttt{str}) -- complete path to save folder
    including filename without extension
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

\hypertarget{classmethod-fetch_meshpubmed_id}{%
\subsubsection{classmethod
fetch\_mesh(pubmed\_id)}\label{classmethod-fetch_meshpubmed_id}}

Fetch MeSH keywords for article based on PubMed ID.

\begin{itemize}
\item
  \textbf{Parameters}

  \textbf{pubmed\_id} (\texttt{str}) -- article PubMed ID
\item
  \textbf{Return type}

  \texttt{List}{[}\texttt{Dict}{]}
\end{itemize}

\hypertarget{classmethod-get_departments}{%
\subsubsection{classmethod
get\_departments()}\label{classmethod-get_departments}}

Get all Edoc departments.

\begin{itemize}
\item
  \textbf{Return type}

  \texttt{List}{[}\texttt{str}{]}
\end{itemize}

\hypertarget{classmethod-inspectdata-fields}{%
\subsubsection{classmethod inspect(data,
*fields)}\label{classmethod-inspectdata-fields}}

Print fields of items in data.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{data} (\texttt{List}{[}\texttt{Dict}{]}) -- data to be
    inspected
  \item
    \textbf{fields} (\texttt{str}) -- fields to focus on
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

\hypertarget{classmethod-map2referencekeyword}{%
\subsubsection{classmethod
map2reference(keyword)}\label{classmethod-map2referencekeyword}}

Map a keyword to its reference keyword.

Keyword must first be cleaned by corresponding \_Keyword method.

\begin{itemize}
\item
  \textbf{Parameters}

  \textbf{keyword} (\texttt{str}) -- the keyword
\item
  \textbf{Return type}

  \texttt{Dict}
\end{itemize}

\hypertarget{classmethod-select_from_datadata-fields}{%
\subsubsection{classmethod select\_from\_data(data,
*fields)}\label{classmethod-select_from_datadata-fields}}

Select items from data based on non-empty fields.

Only items where all required fields are non-empty are sample.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{data} (\texttt{List}{[}\texttt{Dict}{]}) -- the input data
  \item
    \textbf{fields} (\texttt{str}) -- the required fields for an item to
    be sample
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{List}{[}\texttt{Dict}{]}
\end{itemize}

\hypertarget{classmethod-select_from_filefile_path-fields}{%
\subsubsection{classmethod select\_from\_file(file\_path,
*fields)}\label{classmethod-select_from_filefile_path-fields}}

Select items from file according to fields.

For example: select\_from\_file(DIR + ``/raw/2019.json,'' ``title,''
``abstract,'' ``keywords,'' ``id\_number'')

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{file\_path} (\texttt{str}) -- complete path to file
    including filename and extension
  \item
    \textbf{fields} (\texttt{str}) -- the required fields for an item to
    be sample
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

\hypertarget{classmethod-super_enrich_with_annifabstract}{%
\subsubsection{classmethod
super\_enrich\_with\_annif(abstract)}\label{classmethod-super_enrich_with_annifabstract}}

Enrich items with automatic keywords using all Annif-client projects.

\begin{itemize}
\item
  \textbf{Parameters}

  \textbf{abstract} (\texttt{bool}) -- toggle use abstract for indexing
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

\hypertarget{class-files.keywords}{%
\subsection{class files.Keywords()}\label{class-files.keywords}}

A collection of functions for manipulating Edoc author keywords.

\hypertarget{classmethod-clean_keywordkeyword}{%
\subsubsection{classmethod
clean\_keyword(keyword)}\label{classmethod-clean_keywordkeyword}}

Clean a keyword.

\begin{itemize}
\item
  \textbf{Parameters}

  \textbf{keyword} (\texttt{str}) -- the keyword
\item
  \textbf{Return type}

  \texttt{List}{[}\texttt{str}{]}
\end{itemize}

\hypertarget{classmethod-clean_keywordskeywords_per_item}{%
\subsubsection{classmethod
clean\_keywords(keywords\_per\_item)}\label{classmethod-clean_keywordskeywords_per_item}}

Turn a string of keywords per item into a list of cleaned (=normalized)
keywords.

\begin{itemize}
\item
  \textbf{Parameters}

  \textbf{keywords\_per\_item} (\texttt{List}{[}\texttt{str}{]}) -- the
  keywords
\item
  \textbf{Return type}

  \texttt{List}{[}\texttt{str}{]}
\end{itemize}

\hypertarget{classmethod-enrich_with_ysofile_path-save_path}{%
\subsubsection{classmethod enrich\_with\_yso(file\_path,
save\_path)}\label{classmethod-enrich_with_ysofile_path-save_path}}

Blah.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{file\_path} (\texttt{str}) -- complete path to file
    including filename and extension
  \item
    \textbf{save\_path} (\texttt{str}) -- complete path to save folder
    including filename without extension
  \end{itemize}
\end{itemize}

\hypertarget{classmethod-extract_keywordsfile_path}{%
\subsubsection{classmethod
extract\_keywords(file\_path)}\label{classmethod-extract_keywordsfile_path}}

Extract keywords per item from file.

\begin{itemize}
\item
  \textbf{Parameters}

  \textbf{file\_path} (\texttt{str}) -- complete path to file including
  filename and extension
\item
  \textbf{Return type}

  \texttt{List}{[}\texttt{str}{]}
\end{itemize}

\hypertarget{classmethod-fetch_ysokeyword}{%
\subsection{classmethod
fetch\_yso(keyword)}\label{classmethod-fetch_ysokeyword}}

Fetch the YSO ID for a keyword if any.

\begin{itemize}
\item
  \textbf{Parameters}

  \textbf{keyword} (\texttt{str}) -- the keyword
\item
  \textbf{Return type}

  \texttt{Optional}{[}\texttt{str}{]}
\end{itemize}

\hypertarget{classmethod-make_countfile_path-save_path}{%
\subsubsection{classmethod make\_count(file\_path,
save\_path)}\label{classmethod-make_countfile_path-save_path}}

Count all keywords and their respective IDs per item in file.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{file\_path} (\texttt{str}) -- complete path to file
    including filename and extension
  \item
    \textbf{save\_path} (\texttt{str}) -- complete path to save folder
    including filename and extension
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

\hypertarget{classmethod-make_histogramkeywords}{%
\subsubsection{classmethod
make\_histogram(keywords)}\label{classmethod-make_histogramkeywords}}

Make a histogram for keywords.

\begin{itemize}
\item
  \textbf{Parameters}

  \textbf{keywords} (\texttt{List}{[}\texttt{str}{]}) -- the keywords
\item
  \textbf{Return type}

  \texttt{List}{[}\texttt{Dict}{]}
\end{itemize}

\hypertarget{class-files.analysis}{%
\subsection{class files.Analysis()}\label{class-files.analysis}}

A collection of data analysis functions.

\hypertarget{classmethod-count_confusionstandard-suggestions}{%
\subsubsection{classmethod count\_confusion(standard,
suggestions)}\label{classmethod-count_confusionstandard-suggestions}}

Count the values in the confusion matrix for standard and suggestions.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{standard} (\texttt{list}) -- the standard
  \item
    \textbf{suggestions} (\texttt{list}) -- the suggestions
  \end{itemize}
\end{itemize}

\hypertarget{classmethod-extract_standarditem-marker}{%
\subsubsection{classmethod extract\_standard(item,
marker)}\label{classmethod-extract_standarditem-marker}}

Extract gold standard IDs from item.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{item} (\texttt{dict}) -- the Edoc item
  \item
    \textbf{marker} (\texttt{str}) -- the type of ID
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{Optional}{[}\texttt{list}{]}
\end{itemize}

\hypertarget{classmethod-extract_suggestionsitem-marker-n10}{%
\subsubsection{classmethod extract\_suggestions(item, marker,
n=10)}\label{classmethod-extract_suggestionsitem-marker-n10}}

Extract top n Annif suggestion IDs from item.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{item} (\texttt{dict}) -- the Edoc item
  \item
    \textbf{marker} (\texttt{str}) --
    project\_id-abstract-fulltext-limit-threshold
  \item
    \textbf{n} (\texttt{int}) -- number of top IDs (by score) to be
    extracted, defaults to 10
  \end{itemize}
\end{itemize}

\hypertarget{classmethod-get_id_typemarker}{%
\subsubsection{classmethod
get\_id\_type(marker)}\label{classmethod-get_id_typemarker}}

Get the ID type from the project ID.

\begin{itemize}
\item
  \textbf{Parameters}

  \textbf{marker} (\texttt{str}) --
  project\_id-abstract-fulltext-limit-threshold
\item
  \textbf{Return type}

  \texttt{str}
\end{itemize}

\hypertarget{classmethod-get_sklearn_arrayitem-project_id-abstractfalse-fulltextfalse-limitnone-thresholdnone-n10}{%
\subsubsection{classmethod get\_sklearn\_array(item, project\_id,
abstract=False, fulltext=False, limit=None, threshold=None,
n=10)}\label{classmethod-get_sklearn_arrayitem-project_id-abstractfalse-fulltextfalse-limitnone-thresholdnone-n10}}

Get Sklearn y\_true and y\_pred for an item.

Available Annif-client project IDs are yso-en, yso-maui-en,
yso-bonsai-en, yso-fasttext-en, wikidata-en.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{item} (\texttt{dict}) -- sample data set item
  \item
    \textbf{project\_id} (\texttt{str}) -- Annif-client project ID
  \item
    \textbf{abstract} (\texttt{bool}) -- toggle use abstract for
    indexing, defaults to False
  \item
    \textbf{fulltext} (\texttt{bool}) -- toggle use fulltext for
    indexing, defaults to False
  \item
    \textbf{limit} (\texttt{Optional}{[}\texttt{int}{]}) -- Annif-client
    limit, defaults to None
  \item
    \textbf{threshold} (\texttt{Optional}{[}\texttt{int}{]}) --
    Annif-client threshold, defaults to None
  \item
    \textbf{n} (\texttt{int}) -- number of top IDs (by score) to be
    extracted, defaults to 10
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{Optional}{[}\texttt{dict}{]}
\end{itemize}

\hypertarget{classmethod-make_metricsfile_path-project_id-abstractfalse-fulltextfalse-limitnone-thresholdnone-n10-departmentnone}{%
\subsubsection{classmethod make\_metrics(file\_path, project\_id,
abstract=False, fulltext=False, limit=None, threshold=None, n=10,
department=None)}\label{classmethod-make_metricsfile_path-project_id-abstractfalse-fulltextfalse-limitnone-thresholdnone-n10-departmentnone}}

Make Sklearn metrics F1, recall, precision for file.

The output is saved as /metrics/metrics\_\{marker\}.json.

Available Annif-client project IDs are yso-en, yso-maui-en,
yso-bonsai-en, yso-fasttext-en, wikidata-en.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{file\_path} (\texttt{str}) -- complete path to file
    including filename and extension
  \item
    \textbf{project\_id} (\texttt{str}) -- Annif-client project ID
  \item
    \textbf{abstract} (\texttt{bool}) -- toggle use abstract for
    indexing, defaults to False
  \item
    \textbf{fulltext} (\texttt{bool}) -- toggle use fulltext for
    indexing, defaults to False
  \item
    \textbf{limit} (\texttt{Optional}{[}\texttt{int}{]}) -- Annif-client
    limit, defaults to None
  \item
    \textbf{threshold} (\texttt{Optional}{[}\texttt{int}{]}) --
    Annif-client threshold, defaults to None
  \item
    \textbf{n} (\texttt{int}) -- number of top IDs (by score) to be
    extracted per item, defaults to 10
  \item
    \textbf{department} (\texttt{Optional}{[}\texttt{str}{]}) --
    restrict to items from department
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

\hypertarget{classmethod-make_random_samplefile_path-save_path-size}{%
\subsubsection{classmethod make\_random\_sample(file\_path, save\_path,
size)}\label{classmethod-make_random_samplefile_path-save_path-size}}

Save a random sample from the population in the file.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{file\_path} (\texttt{str}) -- complete path to file
    including filename and extension
  \item
    \textbf{save\_path} (\texttt{str}) -- complete path to save folder
    including filename and extension
  \item
    \textbf{size} (\texttt{int}) -- the sample size
  \end{itemize}
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

\hypertarget{classmethod-print_chi_square_fitfile_path}{%
\subsubsection{classmethod
print\_chi\_square\_fit(file\_path)}\label{classmethod-print_chi_square_fitfile_path}}

Print chi square goodness of fit.

File must be CSV with first line title; three columns with first column
categories, second column expected and third column observed values.

\begin{itemize}
\item
  \textbf{Parameters}

  \textbf{file\_path} (\texttt{str}) -- complete path to file including
  filename and extension
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

\hypertarget{classmethod-super_make_metricsfile_path-departmentnone}{%
\subsubsection{classmethod super\_make\_metrics(file\_path,
department=None)}\label{classmethod-super_make_metricsfile_path-departmentnone}}

Make metrics for all combinations of Annif projects and parameters in
enriched Edoc file.

Output files are saved in /metrics/. Joint results are saved in
/analysis/metrics.json.

\begin{itemize}
\item
  \textbf{Parameters}

  \begin{itemize}
  \item
    \textbf{file\_path} (\texttt{str}) -- complete path to file
    including filename and extension
  \item
    \textbf{department} (\texttt{Optional}{[}\texttt{str}{]}) --
    restrict to items from department
  \end{itemize}
\end{itemize}

\hypertarget{classmethod-super_make_stats}{%
\subsubsection{classmethod
super\_make\_stats()}\label{classmethod-super_make_stats}}

Make metrics for files in /metrics.

Output is saved as /analysis/metrics.json.

\begin{itemize}
\item
  \textbf{Return type}

  \texttt{None}
\end{itemize}

>>>>

In this section, the problem and goal of this project are stated and the
employed methodology is outlined.

@@

@@

@@

Even though Edoc is a public server, its database does not have a
web-ready API. In addition, since Edoc is a production server, its
underlying database cannot be used directly on pain of disturbing the
provided services. The data hence needs to be extracted from Edoc in
order to work with it. This could for example be achieved by cloning the
database of the server and running it on a local machine. However, this
route was not available due to the limited resources of the responsible
co-workers. I thus employed a workaround: Edoc has a built-in advanced
search tool where the results can be exported. Even though it is not
possible to extract all database items by default, only one of the many
available search fields has to be filled in. The complete database can
hence be extracted by only filling in the \texttt{date} field and using
a sufficiently permissive time period such as 1940 to 2020 since the
oldest record in the database was published in 1942. On January 21 2021,
this query yielded 68'345 results. These results were then exported as a
326 MB JSON file called \texttt{1942-2020.json}. \texttt{1942-2020.json}
has two drawbacks. First, the maximum file size on GitHub is 100 MB. And
second, \texttt{1942-2020.json} is too big to be handled by standard
editors requiring special editors such as Oxygen. For these reasons
\texttt{1942-2020.json} was split into smaller files that are easier
handled and can be uploaded to GitHub. For this
\texttt{Utility.split\_json} was used yielding 14 files of size 20 MB or
less containing 5000 or fewer entries each. These files are called
\texttt{raw\_master\_\{year\}.json} and saved under \texttt{/raw}.

@@

Taking into account the third constraint simply means that we do not
employ any further restrictions. We can hence construct the sample data
set by choosing exactly those items from \texttt{/raw} which have
non-empty data fields \texttt{title}, \texttt{abstract},
\texttt{id\_number}, and \texttt{keywords}. We do so by calling
\texttt{Data.select\_from\_file} iteratively for all files in
\texttt{/raw}. The resulting file is saved in \texttt{/selected} as
\texttt{selected\_master.json}.

@@

\begin{figure}
\centering
\includegraphics{images/chi_square_selection_fields.pdf}
\caption{The sample data set (\(n\) = 4'111) is not representative of
the raw Edoc data per department. Data field \texttt{abstract}:
\(\chi^2 (df=9) =\) 3'160.556, \(p < 0.001\); data field
\texttt{id\_number}: \(\chi^2 (df=9) =\) 4'209.0285, \(p < 0.001\); data
field \texttt{keywords}: \(\chi^2 (df=9) =\) 2'314.533, \(p < 0.001\).
The data foundation is available at
\texttt{/analysis/chi\_square\_\{data\ field\}} and the
\(\chi^2\)-statistic can be calculated by calling
\texttt{Analysis.print\_chi\_square\_fit} on the data foundation files.}
\end{figure}

Of the 68'345 items in \texttt{/raw}, all have a title (non-empty
\texttt{title} data field), a little more than half of the items have an
abstract (37'381 items with non-empty \texttt{abstract} data field),
roughly half of the items have an ID (35'355 items with non-empty
\texttt{id\_number} data field),\footnote{Note that when using the facet
  by blank on \texttt{id\_number} in OpenRefine, there are 57'153
  matches; this is due to the fact that many items with an ID such as
  DOI have secondary or tertiary IDs such as ISI or PMID.} and less than
10\% of the items have keywords (6'660 items with non-empty
\texttt{keywords} data field). The sample data set as requires all the
above data fields to be non-empty; \texttt{/sample/sample\_master.json}
has 4'111 items and hence constitutes 6\% of the raw data.

@@

\begin{figure}
\centering
\includegraphics{images/raw_sample_analysis.pdf}
\caption{The sample data set (\(n\) = 4'111) is significantly skewed
towards English (data field \texttt{language}: \(\chi^2 (df=3) =\)
1'441.414, \(p < 0.001\)) journal publications (data field
\texttt{type}: \(\chi^2 (df=10) =\) 52'519.743, \(p < 0.001\)) in the
sciences, medicine and economics (data field \texttt{department}:
\(\chi^2 (df=9) =\) 14'447.276, \(p < 0.001\)) from the 21st century
(data field \texttt{date}: \(\chi^2 (df=7) =\) 35'878.493,
\(p < 0.001\)) as compared to the raw Edoc data. The data foundation is
available at \texttt{/analysis/chi\_square\_\{data\ field\}} and the
\(\chi^2\)-statistic can be calculated by calling
\texttt{Analysis.print\_chi\_square\_fit} on the data foundation files.}
\end{figure}

@@

In order to generate the ouput, we call
\texttt{Data.super\_enrich\_with\_annif}, thereby iteratively calling
\texttt{Data.enrich\_with\_annif} for all items in
\texttt{/sample/sample\_master.json} for all Annif projects. The output
is saved in \texttt{/indexed/} as \texttt{indexed\_master.json}.

@@

As explained in \protect\hyperlink{machine-indexing}{section ``Machine
indexing''}, Annif assigns index terms from a controlled vocabulary. If
we want to assess the quality of the indexing via a gold standard, we
must therefore ensure that the gold standard makes use of the vocabulary
used by Annif. The relevant (English) vocabularies are Wikidata and YSO.
The next step in constructing the derivative gold standard is hence to
match the extracted and cleaned keywords with keywords from Wikidata and
YSO. This process is called ``reconciliation'' (see
https://docs.openrefine.org/manual/reconciling) and the tool of choice
for this task is OpenRefine).\footnote{For data refinement I use
  OpenRefine version 3.4.1 (https://openrefine.org/). Data manipulation
  in OpenRefine is tracked: the manipulation history of some data can be
  exported as JSON file and then reproduced (on the same data on a
  different machine or on different data) by loading said file. I will
  supply a corresponding manipulation history whenever appropriate. Note
  that when using OpenRefine with larger files (and there are many such
  files in this project), the memory allocation to OpenRefine must be
  manually increased (see
  https://docs.openrefine.org/manual/installing/\#increasing-memory-allocation
  for details). Also note that OpenRefine always counts the number of
  records by the first column (a fact which caused me many headaches).}

@@

In what follows I will discuss the general results of assessing the
performance of the 100 Annif configurations versus the derivative gold
standard with respect to the Edoc sample data set. The data foundation
is available at \texttt{/analysis/metrics.json}.

@@

A surprising result is that there was no performance gain achieved by
increasing the text basis from titles to titles and abstracts (see
\protect\hyperlink{creating-the-data-foundation}{section ``Creating the
data foundation''}): if anything, the performance for the extended text
basis was slightly worse (see Figure 6, right). There is again no easy
explanation for this find since there are multiple plausible causes
which cannot be distangled here. For example, the abstract of an item in
the Edoc sample data set might be less representative of the item's
content than its title. In that case, the observed worse perfomance of
the Annif configurations based on title and abstracts would even be
expected. Alternatively, it might be the case that the problem lies with
Annif: a larger text basis introduces more confounding factors and might
make it harder to extract the subject terms required. An experimentum
crucis to decide between the two discussed explanations would be to
further extend the text basis to include full texts in addition to
titles and abstracts.

@@

In \protect\hyperlink{general-results}{section ``General results''} I
have identified and discussed the overall best Annif configerations.
However, in \protect\hyperlink{analysis}{section ``Analysis''}, I had
noted the caveat that the performance of Annif might vary according to
department due to systematic biases in constructing the Edoc sample data
set. In this section I will therefore assess the performance of the
various Annif configurations per department.

@@

\hypertarget{conclusion-and-outlook}{%
\section{Conclusion and outlook}\label{conclusion-and-outlook}}

In this section, ..

\hypertarget{recommendation}{%
\subsection{Recommendation}\label{recommendation}}

@@

\hypertarget{further-research}{%
\subsection{Further research}\label{further-research}}

Apart from implementing a production system as discussed above, natural
next steps to build on this project include:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  mesh
\item
  full text
\item
  german vocabulary
\item
  select domain specific vocabularies with bartoc suggest
\end{enumerate}

@@

As mentioned in \protect\hyperlink{introduction}{section
``Introduction''}, all files, data, and code used in this project is
available at https://github.com/MHindermann/mas.
